---
title: "R crash course / 03"
author: ""
date: "`r Sys.Date()`"
output:
  ioslides_presentation: default
---

```{r,echo=FALSE}
## Set default options for the knitr RMD processing
knitr::opts_chunk$set(echo=TRUE,eval=FALSE,warning=FALSE,message=FALSE,fig.width=5,fig.height=5,cache=TRUE,autodep=TRUE, results="hide")
```

# Creating the LABRES data set

## Reading and column selection

```{r}
targets <- read_xlsx("Datasets/expression_data_vaccination_example.xlsx", sheet=1)
targets <- targets[ , c("USUBJID", "Timepoint", "ARM", "SEX", "AGE") ]
lab <- read.csv("Datasets/labresults_full.csv")
lab <- lab[ , c("USUBJID", "LBTEST", "LBTESTCD", "LBORRES", "Timepoint") ]
## not what we want!
merged <- merge(targets, lab, by="USUBJID")
```

## Reading and column selection

We need to choose: do we want to keep all measurements and just fill out
meta-information from `targets`, or do we want only to keep the time points
present in the expression data set?

```{r}
## to keep all time points
targets <- targets[ , c("USUBJID", "ARM", "SEX", "AGE") ]
targets <- targets[ !duplicated(targets$USUBJID), ] ## remove duplicated lines
merged <- merge(targets, lab, by="USUBJID")

## this one drops all timepoints not present in "targets"
merged <- merge(targets, lab, by=c("USUBJID", "Timepoint"))
```

## Converting to wide format

```{r}
## this does not work as expected, because "LBTESTCD" and "LBTEST" contain
## the same information – we cannot store both in the column names
labwide <- spread(merged, key="LBTESTCD", value="LBORRES")

## this works
labwide <- spread(merged[ , -6], key="LBTESTCD", value="LBORRES")
```

## Cleaning up

```{r}
sapply(labwide, class) # all measurements are character!
labwide2 <- data.frame(labwide[,1:5], data.matrix(labwide[,6:36]))
which(is.na(labwide2), arr.ind=T)

## impute low measurements
labwide[,14] <- gsub("<", "", labwide[,14])
labwide[,16] <- gsub("<", "", labwide[,16])

## these column is all NA
labwide <- labwide[,-17]
labwide <- data.frame(labwide[,1:5], data.matrix(labwide[,6:36]))
```


# Simple statistics with R

## Basic tests (demonstration)




## Basic plots (demonstration)
